Parameter Inference\\with Bifurcation Diagrams

### 2021-06-08 docs/appendix.tex
submitting to neurips
arxiv pre-print

\section{Properties of the Bifurcation Measure}
\label{appendix:measure}

@@

% \section{Simplification of two-state model}

% Consider the general two-state model:
% \begin{equation}
%     \dfrac{dv_i}{dt} = \dfrac{a_i + b_i (K_i v_j)^2}{1 + (K_i v_j)^2} - \mu_i v_i
% \end{equation}

% Without loss of generality, we can rescale $v_i = U_i u_i$. Then,
% \begin{equation}
%     \dfrac{U_idu_i}{dt} = \dfrac{a_i + b_i (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i U_i u_i
% \end{equation}

% Dividing through by $U_i$, we obtain
% \begin{equation}
%     \dfrac{du_i}{dt} = \dfrac{\frac{a_i}{U_i} + \frac{b_i}{U_i} (K_i U_j u_j)^2}{1 + (K_i U_j u_j)^2} - \mu_i u_i
% \end{equation}

% By choosing $U_i = b_i$, the model reduces to:
% \begin{equation}
%     \dfrac{du_1}{dt} = \dfrac{\hat{a}_1 + (p u_2)^2}{1 + (p u_2)^2} - \mu_1 u_1, \quad
%     \dfrac{du_2}{dt} = \dfrac{\hat{a}_2 + (k u_1)^2}{1 + (k u_1)^2} - \mu_2 u_2
% \end{equation}
% where $p = \frac{K_1}{\sqrt{b_1}}$ and $k = \frac{K_2}{\sqrt{b_2}}$.

>>>>

\section{Conditions for Bifurcations}
\label{appendix:conditions}

A saddle-node bifurcation is the generic bifurcation of equilibria and additional conditions are required at a bifurcation point to obtain a transcritical or pitchfork bifurcations \cite{Kuznetsov2004TopologicalSystems}. If we are not specifically looking for transcritical or pitchfork bifurcations, it is sufficient to consider a non-zero Hessian determinant and non-zero vector slope
\begin{equation}
    \left|\frac{\partial^2\rates}{\partial u^2}\right| \neq 0\qquad \frac{\partial\rates}{\partial p}\neq 0
    \label{eq:conditions}
\end{equation}
In the multidimensional case we would like to express the above conditions succinctly in terms of a scalar derivative. Let us consider the non-degeneracy condition along the equilibrium manifold in terms of the state-space determinant
\begin{equation}
    \frac{d}{ds}\left|\frac{\partial\rates}{\partial u}\right|\neq 0
\end{equation}
Applying Jacobi's formula and expanding the directional derivative in terms of basis vectors $\hat{u}(s)$ and $\hat{p}(s)$ that are tangent to the equilibrium manifold at a particular location $s$ we get
\begin{equation}
     \hat{p}(s)\,\mathrm{Tr}\left[
        \frac{\partial^2\rates}{\partial u\partial p}
        \,\mathrm{adj}\frac{\partial\rates}{\partial u}
        \right]
        +
    \sum_{n=1}^N\hat{u}_{n}(s)\,\mathrm{Tr}\left[
        \frac{\partial^2\rates}{\partial u \partial u_n}
        \,\mathrm{adj}\frac{\partial F}{\partial u}
        \right]
  \neq 0
\end{equation}
We can see this expression contains the vector Hessian $\frac{\partial^2\rates}{\partial u^2}$ whos determinant is non-zero for the non-degenerate case. The slope $\frac{\partial\rates}{\partial p}$ is non-zero when the transversality condition holds.

@@

### 2021-06-08 docs/article.tex
arxiv pre-print

\begin{itemize}
    \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
    \item Implementation of the method as a Julia package\\  \href{https://github.com/anon/Code.jl}{\texttt{see attached zip file}}
% \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
    \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
\end{itemize}

@@

% \section{Acknowledgements}
% We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach.

@@

\clearpage\pagenumbering{gobble}
\section*{Checklist}
\input{checklist}

>>>>

\begin{itemize}
    \item An end-to-end differentiable method for locating bifurcations in parameter space and then matching their dependency on a control condition to user-specified locations
    \item Implementation of the method as a Julia package\\
\href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}
    \item Leveraging the cost landscape for a novel way of organising differential equation models in terms of geometric and topological equivalence
\end{itemize}

@@

\section{Acknowledgements}
We would like to acknowledge Kieran Cooney for the fruitful conversations that helped guide the derivations and computational approach. A special thanks go to Romain Veltz and the Julia community for helpful pointers on package development and discussions over Slack.

@@

% \clearpage\pagenumbering{gobble}
% \section*{Checklist}
% \input{checklist}
